---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r message=FALSE}
library(dplyr)
```

## Loading and preprocessing the data

```{r}
zipFilename='activity.zip'
dataFilename='activity.csv'
if (!file.exists(dataFilename)) {
	unzip(zipFilename,dataFilename)
}
data=read.csv(dataFilename,colClasses=c('integer','Date','integer'))
data$interval=as.POSIXct(
	paste(
		'2000-01-01',
		paste(
			data$interval%/%100,
			data$interval%%100,
			0,
			sep=':'
		)
	)
)
```


## What is mean total number of steps taken per day?

```{r}
stepsPerDay = (
	data %>% 
	group_by(date) %>% 
	summarize(steps=sum(steps))
)$steps
hist(
	stepsPerDay,
	main='Histogram of total number of steps per day',
	xlab='steps per day'
)
meanStepsPerDay=mean(stepsPerDay,na.rm=T)
medianStepsPerDay=median(stepsPerDay,na.rm=T)
```

Distribution of total steps per day has mean `r format(meanStepsPerDay)` and median `r format(medianStepsPerDay)`.


## What is the average daily activity pattern?

```{r}
dataIntervals = (data %>%
	group_by(interval) %>%
	summarize(steps=mean(steps,na.rm=T))
)
plot(
	dataIntervals,
	type='l',
	main='Average number of steps per interval'
)
maxStepsInterval=dataIntervals[which.max(dataIntervals$steps),]$interval
```

Maximum number of steps is between `r strftime(maxStepsInterval,format="%H:%M")` and `r strftime(maxStepsInterval+5*60,format="%H:%M")`.


## Imputing missing values

```{r}
nMissingValues=sum(is.na(data$steps))
```

Total number of missing values is `r nMissingValues`. They are replaced by mean values for 5-minute intervals.

```{r}
dataImputed=data
dataImputed$steps=ifelse(is.na(data$steps),dataIntervals$steps,data$steps)
```

```{r}
stepsPerDayImputed = (
	dataImputed %>% 
	group_by(date) %>% 
	summarize(steps=sum(steps))
)$steps
hist(
	stepsPerDayImputed,
	main='Histogram of total number of steps per day computed from imputed data',
	xlab='steps per day'
)
meanStepsPerDayImputed=mean(stepsPerDayImputed)
medianStepsPerDayImputed=median(stepsPerDayImputed)
```

Distribution of total steps per day in the imputed data has mean `r format(meanStepsPerDay)` and median `r format(medianStepsPerDay)` -- the same as in the original data. This is not surprising because missing days are replaced by average days.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
dataImputed$dayOfWeekType=factor(ifelse(
	between(as.POSIXlt(dataImputed$date)$wday,1,5),
	'weekday',
	'weekend'
))
dataIntervalsDoWType = (dataImputed %>%
	group_by(interval,dayOfWeekType) %>%
	summarize(steps=mean(steps,na.rm=T))
)
library(ggplot2)
library(scales)
qplot(interval,steps,facets=dayOfWeekType~.,geom='line',data=dataIntervalsDoWType)+scale_x_datetime(labels=date_format('%H:%M'))
```
